Implementation and assessment of tools for integrating forest roads into projections of the cumulative effects of disturbance on wildlife

### 2024-05-13 analysis/paper/supplement.Rmd
more writing in progress

The grade penalty method for calculating edge weights is a simplified version of the approach taken by @anderson2004a that does not distinguish between adverse and favourable grades. We use a modified Digital Elevation Model (DEM) as the input to the edge weight calculation that allows barriers other than slope to be included without incurring additional computation and memory costs. Pixels where roads cannot be built are assigned a value of NA, and negative values indicate costs associated with other features (e.g. stream crossings). Existing roads are assigned a value of 0. All other values in the weight raster are assumed to be elevation and grade is calculated from differences between adjacent locations. Default construction cost values are taken from the BC Interior Appraisal Manual: the base cost of building a road is \$16 178 with an additional cost of \$504 for every 1% grade and a limit of 20% grade. Edges with greater than 20% grade are assigned either NA or an arbitrarily high value. A value of 65 000 is used in our projections to avoid errors when a landing cannot be accessed without crossing an NA cell. Note that grade penalties are not applied adjacent to existing roads, nor do they apply when connecting to a node with a negative value. In the latter case, the edge weight is set to the highest absolute cost.

@@

We compared the speed and peak RAM usage of the MST and ILCP methods with landings regularly sampled at 10 per km^2^ (high density) or using only the cutblock centroids as landings on weight rasters of three different resolutions (1000 m, 500 m and 100 m), giving 18 297, 72 633 and 1 815 825 nodes in the cost graph, respectively. Each cell in the raster becomes a node in the graph used to determine the shortest path which increases the time and memory requirements of the projection. Benchmarking was done using Azure Standard A4 v2 nodes which have 4 cores and 8 GB of RAM, chosen to match a typical laptop. Each projection was run on a separate node and the `peakRAM` package was used to measure the total time to run the projection and the maximum memory used at any point in the analysis [@quinn_peakRAM_2024]. We also timed the execution of the Hardy QGIS plugin with the 100 m resolution weights raster.

>>>>

The grade penalty method for calculating edge weights is a simplified version of the approach taken by Anderson and Nelson [-@anderson_projecting_2004] that does not distinguish between adverse and favourable grades. We use a modified Digital Elevation Model (DEM) as the input to the edge weight calculation that allows barriers other than slope to be included without incurring additional computation and memory costs. Pixels where roads cannot be built are assigned a value of NA, and negative values indicate costs associated with other features (e.g. stream crossings). Existing roads are assigned a value of 0. All other values in the weight raster are assumed to be elevation and grade is calculated from differences between adjacent locations. Default construction cost values are taken from the BC Interior Appraisal Manual: the base cost of building a road is \$16 178 with an additional cost of \$504 for every 1% grade and a limit of 20% grade. Edges with greater than 20% grade are assigned either NA or an arbitrarily high value. A value of 65 000 is used in our projections to avoid errors when a landing cannot be accessed without crossing an NA cell. Note that grade penalties are not applied adjacent to existing roads, nor do they apply when connecting to a node with a negative value. In the latter case, the edge weight is set to the highest absolute cost.

@@

We compared the speed and peak RAM usage of the MST and ILCP methods with landings regularly sampled at 10 per km^2^ (high density) or using only the cutblock centroids as landings on weight rasters of three different resolutions (1000 m, 500 m and 100 m), giving 18 297, 72 633 and 1 815 825 nodes in the cost graph, respectively. Each cell in the raster becomes a node in the graph used to determine the shortest path which increases the time and memory requirements of the projection. Benchmarking was done using Azure Standard A4 v2 nodes which have 4 cores and 8 GB of RAM, chosen to match a typical laptop. Each projection was run on a separate node and the `peakRAM` package was used to measure the total time to run the projection and the maximum memory used at any point in the analysis [@quinn_peakram_2024]. We also timed the execution of the Hardy QGIS plugin with the 100 m resolution weights raster.
